Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-12-17 Paper_draft/main.tex
recent changes 16/11

\acrodef{GW}[GW]{gravitational wave}
\acrodef{CBC}[CBC]{compact binary coalescence} 
\acrodef{ML}[ML]{machine learning}
\acrodef{AI}[AI]{artificial intelligence}
\acrodef{CNN}[CNN]{convolutional neural network}
\acrodef{GAN}[GAN]{generative adversarial network}
\acrodef{CGAN}[CGAN]{conditional generative adversarial network}
\acrodef{ACGAN}[ACGAN]{auxilliary conditional generative adversarial network}
\acrodef{DCGAN}[DCGAN]{deep convolutional generative adversarial network}
\acrodef{CNN}[CNN]{convolutional neural networks}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{PSD}[PSD]{power spectral density}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%textwidth in inches: \printinunitsof{in}\prntlen{\textwidth}
\begin{comment}
\begin{itemize}
\item Need to introduce GWs - the current state of the field e.g. detections
and LVC papers \ding{51}
\item Introduce burst searches - what's the point of burst searches \ding{51} - lots of references 
\item Discuss the family of burst waveforms currently used and why - not in detail, just
an introduction \ding{51}
\item Introduce ML techniques in GWs \ding{51} - lots of references
\item What this paper does on GANs in 1 paragraph \ding{51}
\item Describe the structure of the paper 
\end{itemize}
\end{comment}

% introduce \ac{GW} astrophsyics
%
Gravitational wave astronomy is now an established field that began with the first
detection of a binary black hole merger~\cite{Abbott2016} in September 2015.
Following this, the first and second observations runs (O1 and O2) of Advanced
LIGO and Advanced Virgo~\cite{Prospects-dets, AdvLIGO, AdvLIGO2, AdvVIRGO}
reported several more \ac{CBC} mergers~\cite{Abbott2016a, Abbott2017,
Abbott2017a, Abbott2017b}. On 17th August 2017 a binary neutron star merger was
observed alongside its electromagnetic counterpart for the first time, giving
rise to multi-messenger \ac{GW} astronomy.

@@

% What this paper does on GANs in 1 paragraph - the point of the paper
%
In this work we aim to explore the use of \ac{ML} to generate and interpret
unmodeled \ac{GW} burst waveforms. Using the generative machine learning
model, \acp{GAN}, we train on five classes of waveforms in the time domain. Working on the assumption that \acp{GAN} construct smooth
 high dimensional vector spaces between their input and output, we can then
explore the space between the five classes to construct new
hybrid waveforms. As all the computationally expensive
processes occur during training, once trained, the model is able to
generate waveforms in fractions of a second and produce waveforms that are difficult to generate with current
techniques. These new varieties of waveforms can then be used to diagnose \michael{not sure about diagnose, maybe test or evaluate?}
detection algorithms, gain new insight into sources of \ac{GW}
bursts and  allow us to better train our algorithms on a
broader range of possible signals and therefore enhance our detection ability.

% the structure of the paper
%
This paper is organised as follows. In \cref{ML overview} we introduce the basic ideas of machine learning and discuss the choice of algorithm we used. In \cref{Method} we describe the training data and the details of the model. We present the results of the GAN in \cref{results} and show how unmodeled signals can be produced by interpolating and sampling within latent and class spaces. In \cref{cnn classifier} we show that a convolutional neural network \michael{should this have the accronym? Tried to fix it but it only puts the plural for me} classifier can be trained to distinguish between sets of our GAN generated waveforms from noise only cases. We conclude with a summary of the presented work in \cref{conclusions}.

@@

\begin{figure}[h!]
\begin{subfigure}[b]{0.30\textwidth}
   \begin{subfigure}[b]{1\textwidth}
   	\centering
	\resizebox{\textwidth}{!}{
 	\begin{tikzpicture}		
		    \node[scale=0.3] 		   (x1) at (0,0.8) {$x_1$}; 
		    \node[scale=0.3] 		   (x2) at (0,0.5) {$x_2$};  
		    \node[scale=0.3] 		   (x3) at (0,0.2) {$x_3$};       
    		    \node [draw, circle,scale=0.18,fill=blue!25]          (c3) at (0.7,0.5) {\Large $\sigma(\sum_i w_i x_i + b)$};
    \draw[->] (x1)--(c3);
    \draw[->] (x2)--(c3);
    \draw[->] (x3)--(c3);
     \draw[->] (c3)--(1.3,0.5);
   	\end{tikzpicture}
	}
	\caption{}
	 \label{fig:perceptron}
 \end{subfigure}

   \begin{subfigure}[b]{1\textwidth}
  	\centering
	\resizebox{\textwidth}{!}{
\begin{tikzpicture}
\begin{axis}[
    xmin=-2.5, xmax=2.5,
    ymin=-1.5, ymax=1.5,
    axis lines=center,
    axis on top=true,
    domain=-2.5:2.5,
    ylabel=$\sigma$,
    xlabel=$x$,
    ]

    \addplot [mark=none,draw=purple,domain=-2.5:2.5,ultra thick] {1/(1+exp(-x))};
    \addplot+[mark=none,draw=green,domain=-2.5:0,ultra thick] {0};
    \addplot+[mark=none,draw=green,domain=0:2.5,ultra thick] {x};
    \addplot+[mark=none,orange,domain=-2.5:0,ultra thick] {0.2*x};
    \addplot+[mark=none,orange,domain=0:2.5,ultra thick] {x};

    %% Add the asymptotes
\end{axis}
\end{tikzpicture}
	}
	\caption{}
	 \label{fig:activations}
\end{subfigure}
\end{subfigure}
	 \begin{subfigure}[b]{0.70\textwidth}
\centering
\resizebox{\textwidth}{!}{
\def\layersep{1.5cm}
\begin{tikzpicture}[shorten >=1pt,draw=black, ->,node distance=\layersep]

    \tikzstyle{every pin edge}=[<-,shorten <=1pt]
    \tikzstyle{neuron}=[circle,draw=black, very thick, fill=blue!25, minimum size=22pt,inner sep=0pt]
    \tikzstyle{input neuron}=[circle,draw=black,very thick,fill=black!25, minimum size=22pt,inner sep=0pt];
    \tikzstyle{output neuron}=[neuron];
    \tikzstyle{hidden neuron}=[circle,draw=black,very thick,fill=blue!25,minimum size=22pt,inner sep=0pt];
    \tikzstyle{annot} = [text width=4em, text centered]

    % Draw the input layer nodes
    \foreach \name / \y in {1,...,5}
    % This is the same as writing \foreach \name / \y in {1/1,2/2,3/3,4/4}
        \node[input neuron] (I-\name) at (0,-\y) {$x_{\name}$};

    % Draw the hidden layer nodes
    \foreach \name / \y in {1,...,3}
        \path[yshift=-1cm]
            node[hidden neuron] (H-\name) at (\layersep,-\y cm) {};

  \foreach \namee / \y in {1,...,3}
        \path[yshift=-1 cm]
            node[hidden neuron] (H1-\namee) at (2*\layersep,-\y cm){};

    % Draw the output layer node
  \node[output neuron,pin={[pin edge={->}]right:}, right of=H1-2] (O) {};
    % Connect every node in the input layer with every node in the
    % hidden layer.
    \foreach \source in {1,...,5}
        \foreach \dest in {1,...,3}
            \path (I-\source) edge (H-\dest);

 \foreach \source in {1,...,3}
        \foreach \dest in {1,...,3}
            \path (H-\source) edge (H1-\dest);

    % Connect every node in the hidden layer with the output layer
    \foreach \source in {1,...,3}
        \path (H1-\source) edge (O);

    % Annotate the layers
    %\node[annot,above of=H-1, node distance=1cm] (hl) {Hidden layers};
   % \node[annot,left of=I-4] at (0) {Input layer};
    %\node[annot,right of=hl] {Output layer};

    \draw[decoration={brace,mirror,raise=15pt},decorate,-]
  (I-1.north) -- node[left=0.8cm,scale=0.7] {Input layer} (I-5.south);

    \draw[decoration={brace,raise=15pt},decorate,-]
  (H-1.west) --  node[above=0.8cm,scale=0.7]{Hidden layers}  (H1-1.east);

    \draw[decoration={brace,raise=15pt},decorate,-]
  (O.west) --  node[above=0.8cm,scale=0.7]{Output layers}  (O.east);

\end{tikzpicture}
}
 \caption{}
 \label{fig:network}
 \end{subfigure}

\caption{Neural Networks (a) A single neuron taking a vector of inputs and
returning a single output based on the weights, bias and activation function
of the network. (b) A selection of activation functions used in this study. The sigmoid (purple), rectified linear unit (ReLU) \cite{relu} (orange) and leaky rectified linear unit \cite{Maas2013RectifierNI} (green). (c)
A an example of a neural network containing two hidden layers that performs a
mapping of an input vector to a single output. \michael{Minor comment, it's hard to see what the ReLU is doing above 0, maybe use different line styles for each?}}
\end{figure}
%
% Introduce basic neural networks - a perceptron layer
%
\begin{comment}
Neural networks are the quintessential~\chris{really? quintessential?} \jordan{It's a perfectly cromulent word https://www.imdb.com/title/tt0701155/} \ac{ML}
algorithm that aims to approximate a function. 
\end{comment}

\ac{ML} algorithms aim to learn apparent relationships held within given data or `training
data' in order to make accurate predictions without the need for additional
programming. A common approach in \ac{ML} relies on the model learning from
past experience to make decisions on future events. Artificial neural networks are universal function approximators that are built from many single
processing units called neurons. The simplest neural network is the perceptron
layer~\cref{fig:perceptron} which shows a single neuron that takes a vector of real
inputs $x_{i},\ldots, x_{n}$
and maps them to an output according to the linear function, 
%
\begin{align}
f(x) = \sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{align}
%
where $w$ and $b$ are the weights and bias and $\sigma$
denotes the activation function. The weights are numbers which can be thought
of as the strength between connected neurons. The output of a neuron is defined by its activation function which controls how the neuron `fires' depending on its input. Some examples of commonly used activation functions are shown in~\cref{fig:activations}. It is often useful to introduce a bias, $b$, such that the neuron remains inactive
above zero but is active when the sum reaches a defined threshold.

@@

As \acp{GAN} are trained by updating one model at the expense of the other, they can be hard to train. GANs attempting to replicate complicated structures that do not
have the necessary architecture either struggle to produce results at all or
fall into the common failure mode know as mode collapse; where the generator
produces a small variety of samples or simply memorises the training set. The goal of \ac{GAN} training is to find an equilibrium between the two models, if this cannot be found then it is said that the \ac{GAN} has failed to converge. One way to diagnose \acp{GAN} during the development process is to keep track of the loss and accuracy over time \michael{what is being diagnosed? Is development the same as training? I get what this sentence is saying but I think it could be re-written}. Loss plots, for example, as seen in \cref{fig:lossplot} can help to identify common failure modes or to check if the \ac{GAN} has indeed converged. Accuracy is another metric that may be used to monitor convergence and is defined as the number of correct predictions made divided by total number of predictions. There is currently no notion of early stopping in \acp{GAN}, instead, training is halted after convergence and by visually inspecting the generations. 
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/losses.png}
    \caption{Plot of the discriminator and generator loss and accuracy as a function of epochs. Early in training the losses oscillate as both models attempt to find their equilibrium, after which, both losses vary around a point which signifies stable training. Accuracies on the real and fake data are similar, showing that neither model is stronger than the other.}
    \label{fig:lossplot}
\end{figure}

@@

% list the 5 waveform classes
%
\begin{itemize}
%
\item {\bf Sine-Gaussian}: $h_{\text{sg}}(t) = A \exp\left[ - (t-t_{0})^2 /
\tau^2 \right] \sin (2 \pi f_0 (t-t_0) + \phi)$, \michael{Could capitalise SG here?} a sinusoidal wave with a Gaussian
envelope characterised by a central frequency, $f_0$, amplitude, $A$, time of arrival, $t_{0}$ and phase, $\phi$ which is uniformly sampled between $[0, 2\pi]$ \michael{The commas here are a bit of mess, I find it hard to read I think either remove the comma before each symbol or use brackets? The same applies to this whole list. I've removed them in the next point, see what you think.}. 
%
\item {\bf Ringdown}: $h_{\text{rd}}(t) = A \exp \left[-{(t-t_0)} / {\tau}
\right] \sin(2 \pi f_0 (t-t_0) + \phi)$, with frequency $f_0$ and duration $\tau$, amplitude $A$, time of arrival $t_{0}$ and phase $\phi$ which is uniformly sampled between $[0, 2\pi]$. 
%
\item {\bf White noise bursts}: $h_{\text{wn}}(t_j) = Ag_j\exp\left[ -
(t-t_{0})^2 / \tau^2 \right]$ where $g_j$ are drawn from a zero mean unit
variance Gaussian distribution with a Gaussian envelope of duration $tau$.
%
\item {\bf Gaussian pulse}: $h_{\text{gp}}(t) = \exp(-t^2 / \tau^2)$ with
duration parameter $\tau$.
%
\item {\bf Binary black hole}: Simulated using the IMRPhenomD
waveform~\cite{Khan_2016} routine from LALSuite~\cite{lalsuite} which models
the inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses
lie in the range of [30,70] $\textrm{M}_{\odot}$ with zero spins and we fix
$m_1>m_2$. The mass distribution is approximated by a power law with
index of 1.6~\cite{Abbott_2019}. The inclinations are drawn
such that the cosine of the angles lies uniformly in the range [-1,1] and we only use the plus polarisation.
%
\end{itemize}
%
The location of the peak amplitude of the waveforms (corresponding to the
mid-points of all but the ringdown and \ac{BBH} classes) are randomly drawn from a uniform distribution to
be within $[0.4, 0.6]$ seconds from the start of the 1 sec time interval and all
training waveforms are sampled at 1024 Hz.  The parameter prior ranges are
defined in~\cref{Tab:training_parms} and a sample of training waveforms are shown in \cref{fig:training_waveforms}. All training data is rescaled such that their amplitudes peak at 1.

\begin{table}[!h]
\centering
\caption{The parameters used in generating the training data. Each parameter is drawn uniformly in the below ranges. \michael{You've used a mix of - and N/A, are they the same thing or do they indicate something different?} \jordan{I changed N/A to "-" as they do really mean the same thing.}}
%\footnotesize
\begin{tabular}{@{} l l l l l l }
\br
\hline
 Waveform & Central frequency  & Decay & Central time epoch & Mass range \\
 & (Hz) & (s) & (s) & ($\textrm{M}_{\odot}$) \\
\mr
Sine-Gaussian & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & -  \\  
Ringdown & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & - \\
white noise burst & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & -  \\
Gaussian pulse & - & 0.004 - 0.03 & 0.4 - 0.6 & -  \\
BBH & - & - & - & 30 - 70  \\
 \br
\end{tabular}\\
\label{Tab:training_parms}
\end{table}
\normalsize

@@

All models were designed with the Python Keras library \cite{} and TensorFlow \cite{} \michael{Just a reminder these are missing} and trained on a GeForce RTX 2080 Ti GPU. We train the networks for 500 epochs which takes $\mathcal{O}(10)$ hours and save the model at each epoch. We choose an appropriate model by visually inspecting the generations at a point of convergence on the loss plot. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results} \label{results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

While the \ac{GAN} is trained on distinct one-hot encoded classes, we may test arbitrary points in the \ndimensional{5} class space to produce indistinct or hybrid waveforms. In order to explore the class space, in \cref{fig:c_interp} we show results where the latent vector is held constant but we instead linearly
interpolate within the one-hot encoded class space between pairs of the well-defined training class locations. In this scenario we highlight that the \ac{GAN} has not yet probed this intermediate class space during its training and therefore we are reliant on the generator having learned any underlying class space relationships between the 5 training classes. The results show that for each case that the generated signals show distinct characteristics of the respective class pairs at most stages of the transition. We note that transitions in some cases appear to be rather abrupt, e.g., between the Gaussian pulse and the \ac{BBH}, \chris{but we point out that this feature, whilst not uncommon, is a strong function of the random latent space location.} \michael{We sometimes hyphenate class space and sometimes don't}
%
\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/generations/sg-rd.png}
    \includegraphics[width=\textwidth]{figures/generations/rd-wnb.png}
    \includegraphics[width=\textwidth]{figures/generations/wnb-blip.png}
    \includegraphics[width=\textwidth]{figures/generations/blip-bbh.png}
    \caption{\ac{GAN} generated class interpolated waveforms as a function of time showing class space interpolations. A single latent space vector is used for all generations and is chosen randomly in the latent space. Each row shows generations using linearly interpolated classes as inputs to the generator. By row top to bottom: Sine-Gaussian to ringdown, ringdown to white noise burst, Gaussian pulse to BBH.\chris{add class vectors on each panel}}
    \label{fig:c_interp}
\end{figure}
%
\subsection{General points within the class space}
We have shown that the \acp{GAN} latent space and class space \michael{Same comment about hyphenation for latent space.} have structure that can be navigated via interpolation between pairs of locations within each respective space. Taking a step further, we can sample from the class space in novel ways to create new inputs for the generator. These new points are categorised by the method used to sample from the class space. The methods we use are divided into the following: 
%
\begin{itemize}
%
\item {\bf Vertex}: Points that lie at the corners of the \ndimensional{5} class space. These class space locations are equivalent to the training set locations and are our closest generated representation of the training set.
%
\item {\bf Simplex}: This class vector we define as uniformly sampled points on a simplex, which is a generalization of a triangle in $k$-dimensions. We sample uniformly on the $k=4$ simplex that is embedded in the \ndimensional{5} class hyper-cube. In practice we use the equivalent of sampling points from a $k=4$ Dirichlet distribution. It is useful to think of the simplex as the hyper-plane that intersects all 5 training classes. It is a subspace of the Uniform method.  
%
\item {\bf Uniform}: Each of the entries in the class vector is sampled from a uniform distribution $\text{U}[0,1]$. This is equivalent to sampling uniformly within the \ndimensional{5} one-hot encoding hyper-cube.
%
\end{itemize}

@@

\subsection{CNN architecture}
In this approach the input to the \ac{CNN} are 1024 sample time-series (with 2 channels representing each detector output) which are passed through a series of 4 convolutional layers, onto two fully connected or ``Dense'' layers and finally to a single output neuron which represents the probability that a signal is present within the noise. We used dropout in the final dense layer and used a selection of different activation functions including the swish activation \cite{ramachandran2017searching} which improved overall performance and a sigmoid activation for the output layer. We used binary cross-entropy \cref{eqn:crossentropy} as the loss function and Adam as an optimizer with learning rate set to $10^{-3}$. In total we train three separate \acp{CNN} on the vertex, uniform and simplex datasets respectively. In each case the networks share the same architecture and hyperparameters which are defined in \cref{Tab:cnn_training_parms}.

>>>>

\acrodef{GW}[GW]{gravitational wave}
\acrodef{CBC}[CBC]{compact binary coalescence} 
\acrodef{ML}[ML]{machine learning}
\acrodef{AI}[AI]{artificial intelligence}
\acrodef{CNN}[CNN]{convolutional neural network}
\acrodef{GAN}[GAN]{generative adversarial network}
\acrodef{CGAN}[CGAN]{conditional generative adversarial network}
\acrodef{ACGAN}[ACGAN]{auxilliary conditional generative adversarial network}
\acrodef{DCGAN}[DCGAN]{deep convolutional generative adversarial network}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{PSD}[PSD]{power spectral density}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
textwidth in inches: \printinunitsof{in}\prntlen{\textwidth}
\begin{comment}
\begin{itemize}
\item Need to introduce GWs - the current state of the field e.g. detections
and LVC papers \ding{51}
\item Introduce burst searches - what's the point of burst searches \ding{51} - lots of references 
\item Discuss the family of burst waveforms currently used and why - not in detail, just
an introduction \ding{51}
\item Introduce ML techniques in GWs \ding{51} - lots of references
\item What this paper does on GANs in 1 paragraph \ding{51}
\item Describe the structure of the paper 
\end{itemize}
\end{comment}

% introduce \ac{GW} astrophsyics
%
\Ac{GW} astronomy is now an established field that began with the first
detection of a binary black hole merger~\cite{Abbott2016} in September 2015.
Following this, the first and second observations runs (O1 and O2) of Advanced
LIGO and Advanced Virgo~\cite{Prospects-dets, AdvLIGO, AdvLIGO2, AdvVIRGO}
reported several more \ac{CBC} mergers~\cite{Abbott2016a, Abbott2017,
Abbott2017a, Abbott2017b}. On 17th August 2017 a binary neutron star merger was
observed alongside its electromagnetic counterpart for the first time, giving
rise to multi-messenger \ac{GW} astronomy.

@@

% What this paper does on GANs in 1 paragraph - the point of the paper
%
In this work we aim to explore the use of \ac{ML} to generate and interpret
unmodeled \ac{GW} burst waveforms. Using the generative machine learning
model, \acp{GAN}, we train on five classes of waveforms in the time domain. Working on the assumption that \acp{GAN} construct smooth
 high dimensional vector spaces between their input and output, we can then
explore the space between the five classes to construct new
hybrid waveforms. As all the computationally expensive
processes occur during training, once trained, the model is able to
generate waveforms in fractions of a second and produce waveforms that are difficult to generate with current
techniques. These new varieties of waveforms can then be used to evaluate
detection algorithms, gain new insight into sources of \ac{GW}
bursts and  allow us to better train our algorithms on a
broader range of possible signals and therefore enhance our detection ability.

% the structure of the paper
%
This paper is organised as follows. In \cref{ML overview} we introduce the basic ideas of machine learning and discuss the choice of algorithm we used. In \cref{Method} we describe the training data and the details of the model. We present the results of the GAN in \cref{results} and show how unmodeled signals can be produced by interpolating and sampling within latent and class spaces. In \cref{cnn classifier} we show that a \ac{CNN} classifier can be trained to distinguish between sets of our \ac{GAN} generated waveforms from noise only cases. We conclude with a summary of the presented work in \cref{conclusions}.

@@

\begin{figure}[h!]
\begin{subfigure}[b]{0.30\textwidth}
   \begin{subfigure}[b]{1\textwidth}
   	\centering
	\resizebox{\textwidth}{!}{
 	\includegraphics{figures/neuron.pdf}
	}
	\caption{}
	 \label{fig:perceptron}
 \end{subfigure}

   \begin{subfigure}[b]{1\textwidth}
  	\centering
	\resizebox{\textwidth}{!}{
\includegraphics{figures/activation.pdf}}
\caption{}
\label{fig:activation}
\end{subfigure}
\end{subfigure}
	 \begin{subfigure}[b]{0.70\textwidth}
\centering
\resizebox{\textwidth}{!}{

\includegraphics{figures/network.pdf}
}
 \caption{}
 \label{fig:network}
 \end{subfigure}

\caption{Neural Networks (a) A single neuron taking a vector of inputs and
returning a single output based on the weights, bias and activation function
of the network. (b) A selection of activation functions used in this study. The sigmoid (solid purple), rectified linear unit (ReLU) \cite{relu} (dashed wild strawberry red) and leaky rectified linear unit \cite{Maas2013RectifierNI} (dotted blue). (c)
A an example of a neural network containing two hidden layers that performs a
mapping of an input vector to a single output.}
\end{figure}
%
% Introduce basic neural networks - a perceptron layer
%
\begin{comment}
Neural networks are the quintessential~\chris{really? quintessential?} \jordan{It's a perfectly cromulent word https://www.imdb.com/title/tt0701155/} \ac{ML}
algorithm that aims to approximate a function. 
\end{comment}

\ac{ML} algorithms aim to learn apparent relationships held within given data or `training
data' in order to make accurate predictions without the need for additional
programming. A common approach in \ac{ML} relies on the model learning from
past experience to make decisions on future events. Artificial neural networks are universal function approximators that are built from many single
processing units called neurons. The simplest neural network is the perceptron
layer~\cref{fig:perceptron} which shows a single neuron that takes a vector of real
inputs $x_{i},\ldots, x_{n}$
and maps them to an output according to the linear function, 
%
\begin{align}
f(x) = \sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{align}
%
where $w$ and $b$ are the weights and bias and $\sigma$ denotes the activation function. The weights are numbers which can be thought of as the strength between connected neurons. The output of a neuron is defined by its activation function which controls how the neuron `fires' depending on its input. Some examples of commonly used activation functions are shown in~\cref{fig:activation}. It is often useful to introduce a bias, $b$, such that the neuron remains inactive above zero but is active when the sum reaches a defined threshold.

@@

As \acp{GAN} are trained by updating one model at the expense of the other, they can be hard to train. GANs attempting to replicate complicated structures that do not
have the necessary architecture either struggle to produce results at all or
fall into the common failure mode know as mode collapse; where the generator
produces a small variety of samples or simply memorises the training set. The goal of \ac{GAN} training is to find an equilibrium between the two models, if this cannot be found then it is said that the \ac{GAN} has failed to converge. One way to diagnose problems, such as mode collapse, when training \acp{GAN} is to keep track of the loss and accuracy over time. Loss plots, for example, as seen in \cref{fig:lossplot} can help to identify common failure modes or to check if the \ac{GAN} has indeed converged. Accuracy is another metric that may be used to monitor convergence and is defined as the number of correct predictions made divided by total number of predictions. There is currently no notion of early stopping in \acp{GAN}, instead, training is halted after convergence and by visually inspecting the generations. 
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/losses.png}
    \caption{Plot of the discriminator and generator loss and accuracy as a function of epochs. Early in training the losses oscillate as both models attempt to find their equilibrium, after which, both losses vary around a point which signifies stable training. Accuracies on the real and fake data are similar, showing that neither model is stronger than the other.}
    \label{fig:lossplot}
\end{figure}

@@

% list the 5 waveform classes
%
\begin{itemize}
%
\item {\bf Sine-Gaussian}: $h_{\text{SG}}(t) = A \exp\left[ - (t-t_{0})^2 /
\tau^2 \right] \sin (2 \pi f_0 (t-t_0) + \phi)$, \michael{Could capitalise SG here?} a sinusoidal wave with a Gaussian
envelope characterised by a central frequency $f_0$, amplitude $A$, time of arrival $t_{0}$ and phase $\phi$ which is uniformly sampled between $[0, 2\pi]$.
%
\item {\bf Ringdown}: $h_{\text{RD}}(t) = A \exp \left[-{(t-t_0)} / {\tau}
\right] \sin(2 \pi f_0 (t-t_0) + \phi)$, with frequency $f_0$ and duration $\tau$, amplitude $A$, time of arrival $t_{0}$ and phase $\phi$ which is uniformly sampled between $[0, 2\pi]$. 
%
\item {\bf White noise bursts}: $h_{\text{WN}}(t_j) = Ag_j\exp\left[ -
(t-t_{0})^2 / \tau^2 \right]$ where $g_j$ are drawn from a zero mean unit
variance Gaussian distribution with a Gaussian envelope of duration $tau$.
%
\item {\bf Gaussian pulse}: $h_{\text{GP}}(t) = \exp(-t^2 / \tau^2)$ with
duration parameter $\tau$.
%
\item {\bf Binary black hole}: Simulated using the IMRPhenomD
waveform~\cite{Khan_2016} routine from LALSuite~\cite{lalsuite} which models
the inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses
lie in the range of [30,70] $\textrm{M}_{\odot}$ with zero spins and we fix
$m_1>m_2$. The mass distribution is approximated by a power law with
index of 1.6~\cite{Abbott_2019}. The inclinations are drawn
such that the cosine of the angles lies uniformly in the range [-1,1] and we only use the plus polarisation.
%
\end{itemize}
%
The location of the peak amplitude of the waveforms (corresponding to the
mid-points of all but the ringdown and \ac{BBH} classes) are randomly drawn from a uniform distribution to
be within $[0.4, 0.6]$ seconds from the start of the 1 sec time interval and all
training waveforms are sampled at 1024 Hz.  The parameter prior ranges are
defined in~\cref{Tab:training_parms} and a sample of training waveforms are shown in \cref{fig:training_waveforms}. All training data is rescaled such that their amplitudes peak at 1.

\begin{table}[!h]
\centering
\caption{The parameters used in generating the training data. Each parameter is drawn uniformly in the below ranges.}
%\footnotesize
\begin{tabular}{@{} l l l l l l }
\br
\hline
 Waveform & Central frequency  & Decay & Central time epoch & Mass range \\
 & (Hz) & (s) & (s) & ($\textrm{M}_{\odot}$) \\
\mr
Sine-Gaussian & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & -  \\  
Ringdown & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & - \\
white noise burst & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & -  \\
Gaussian pulse & - & 0.004 - 0.03 & 0.4 - 0.6 & -  \\
BBH & - & - & - & 30 - 70  \\
 \br
\end{tabular}\\
\label{Tab:training_parms}
\end{table}
\normalsize

@@

All models were designed with the Python Keras library \cite{chollet2015keras} and TensorFlow \cite{tensorflow2015-whitepaper} and trained on a GeForce RTX 2080 Ti GPU. We train the networks for 500 epochs which takes $\mathcal{O}(10)$ hours and save the model at each epoch. We choose an appropriate model by visually inspecting the generations at a point of convergence on the loss plot. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results} \label{results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

While the \ac{GAN} is trained on distinct one-hot encoded classes, we may test arbitrary points in the \ndimensional{5} class space to produce indistinct or hybrid waveforms. In order to explore the class space, in \cref{fig:c_interp} we show results where the latent vector is held constant but we instead linearly
interpolate within the one-hot encoded class space between pairs of the well-defined training class locations. In this scenario we highlight that the \ac{GAN} has not yet probed this intermediate class space during its training and therefore we are reliant on the generator having learned any underlying class space relationships between the 5 training classes. The results show that for each case that the generated signals show distinct characteristics of the respective class pairs at most stages of the transition. We note that transitions in some cases appear to be rather abrupt, e.g., between the Gaussian pulse and the \ac{BBH}, \chris{but we point out that this feature, whilst not uncommon, is a strong function of the random latent space location.} 
%
\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/generations/sg-rd.png}
    \includegraphics[width=\textwidth]{figures/generations/rd-wnb.png}
    \includegraphics[width=\textwidth]{figures/generations/wnb-blip.png}
    \includegraphics[width=\textwidth]{figures/generations/blip-bbh.png}
    \caption{\ac{GAN} generated class interpolated waveforms as a function of time showing class space interpolations. A single latent space vector is used for all generations and is chosen randomly in the latent space. Each row shows generations using linearly interpolated classes as inputs to the generator. By row top to bottom: Sine-Gaussian to ringdown, ringdown to white noise burst, Gaussian pulse to BBH.\chris{add class vectors on each panel}}
    \label{fig:c_interp}
\end{figure}
%
\subsection{General points within the class space}
We have shown that the \acp{GAN} latent space and class space have structure that can be navigated via interpolation between pairs of locations within each respective space. Taking a step further, we can sample from the class space in novel ways to create new inputs for the generator. These new points are categorised by the method used to sample from the class space. The methods we use are divided into the following: 
%
\begin{itemize}
%
\item {\bf Vertex}: Points that lie at the corners of the \ndimensional{5} class space. These class space locations are equivalent to the training set locations and are our closest generated representation of the training set.
%
\item {\bf Simplex}: This class vector we define as uniformly sampled points on a simplex, which is a generalization of a triangle in $k$-dimensions. We sample uniformly on the $k=4$ simplex that is embedded in the \ndimensional{5} class hyper-cube. In practice we use the equivalent of sampling points from a $k=4$ Dirichlet distribution. It is useful to think of the simplex as the hyper-plane that intersects all 5 training classes. It is a subspace of the Uniform method.  
%
\item {\bf Uniform}: Each of the entries in the class vector is sampled from a uniform distribution $\text{U}[0,1]$. This is equivalent to sampling uniformly within the \ndimensional{5} one-hot encoding hyper-cube.
%
\end{itemize}

@@

\subsection{CNN architecture}
In this approach the input to the \ac{CNN} are 1024 sample time-series (with 2 channels representing each detector output) which are passed through a series of 4 convolutional layers, onto two fully connected or ``Dense'' layers and finally to a single output neuron which represents the probability that a signal is present within the noise. We used dropout in the final dense layer and used a selection of different activation functions including the swish activation \cite{ramachandran2017searching} which improved overall performance, and a sigmoid activation for the output layer. We used binary cross-entropy \cref{eqn:crossentropy} as the loss function and Adam as an optimizer with learning rate set to $10^{-3}$. In total we train three separate \acp{CNN} on the vertex, uniform and simplex datasets respectively. In each case the networks share the same architecture and hyperparameters which are defined in \cref{Tab:cnn_training_parms}.

\begin{figure}[h!] % "[t!]" placement specifier just for this example
\centering
\begin{subfigure}{1\textwidth}
\centering
\includegraphics[width=0.7\linewidth]{figures/efficiency_curve_vtest.png}
%\includegraphics[]{Paper_draft/figures/test.pdf}
\caption{Tested on vertex}
\end{subfigure}
